---
title: "Sym_uniform vs Normal"
author: "David Gerard"
date: "May 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

The difference between $t$ and $normal$ might just be the difference between using a mixture of symmetric uniforms and using a mixture of normals.

Here's an example where the difference is big

```{r}
set.seed(8)
library(tidyverse)
mat <- t(as.matrix(read.csv("../../reproduce_ruv3/Output/gtex_tissue_gene_reads_v6p/muscle.csv",
                            header = TRUE)[, -c(1,2)]))
sout <- seqgendiff::poisthin(mat = mat, nsamp = 6, ngene = 1000,
                             gselect = "mean_max", 
                             prop_null = 0.5)

Y <- log2(sout$Y + 1)

num_sv <- sva::num.sv(dat = t(Y), mod = sout$X)

mout_n <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv)
mout_su <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, mixing_dist = "sym_uniform", 
                            likelihood = "normal")
mout_n$pi0
mout_su$pi0
```

I think the normal is relatively stable with regards to the grid chosen. But the uniform is not. This is way the $t$ probably works worse --- the uniform grid is just unstable sometimes for some reason. I will demonstrate this by choosing a bunch of different grids.

```{r}
dense_grid <- c(0, 2 ^ seq(-40, 40, length = 80))
mout_nd <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, grid_seq = dense_grid)
mout_sud <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, mixing_dist = "sym_uniform", 
                            likelihood = "normal", grid_seq = dense_grid)
mout_nd$pi0
mout_sud$pi0

dense_grid <- c(0, 2 ^ seq(-30, 30, length = 60))
mout_nd <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, grid_seq = dense_grid)
mout_sud <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, mixing_dist = "sym_uniform", 
                            likelihood = "normal", grid_seq = dense_grid)
mout_nd$pi0
mout_sud$pi0

dense_grid <- c(0, 2 ^ seq(-20, 20, length = 40))
mout_nd <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, grid_seq = dense_grid)
mout_sud <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, mixing_dist = "sym_uniform", 
                            likelihood = "normal", grid_seq = dense_grid)
mout_nd$pi0
mout_sud$pi0

dense_grid <- c(0, 2 ^ seq(-20, 10, length = 30))
mout_nd <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, grid_seq = dense_grid)
mout_sud <- vicar::mouthwash(Y = Y, X = sout$X, k = num_sv, mixing_dist = "sym_uniform", 
                            likelihood = "normal", grid_seq = dense_grid)
mout_nd$pi0
mout_sud$pi0
```

```{r}
sessionInfo()
```
