---
title: "Make Sure c Calibration is Correct"
author: "David Gerard"
date: "April 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract
Here, I double check my control-gene based variance inflation calibration code I used during my simulations.

# Analysis

First, generate simulated data
```{r}
library(tidyverse)
## Load code I used in sims
source("../../reproduce_ruv3/Code/nc_adjustment_methods.R")
set.seed(7)
n <- 20
p <- 100
k <- 3
q <- 2

X <- matrix(rnorm(n * q), nrow = n)
beta <- matrix(rnorm(q * p), nrow = q)
beta[, 1:25] <- 0
Z <- matrix(rnorm(n * k), nrow = n)
alpha <- matrix(rnorm(k * p), nrow = k)
E <- matrix(rnorm(n * p), nrow = n)
Y <- X %*% beta + Z %*% alpha + E
ctl <- rep(FALSE, length = p)
ctl[1:13] <- TRUE


vout <- vicar::vruv4(Y = Y, X = X, ctl = ctl, k = k, cov_of_interest = 2, likelihood = "normal",
                     limmashrink = FALSE, include_intercept = FALSE, gls = FALSE)
obj <- list()
obj$betahat   <- vout$betahat
obj$sebetahat <- vout$sebetahat_ols
obj$df        <- nrow(Y) - ncol(X) - k

ctl_out <- ctl_adjust(obj, control_genes = ctl)



dat <- data_frame(SE1 = c(ctl_out$sebetahat), SE2 = c(vout$sebetahat))
ggplot(data = dat, mapping = aes(x = SE1, y = SE2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_bw() +
  xlab("GB SE") +
  ylab("My SE") +
  ggtitle("My SE vs GB SE")

vout$sebetahat[1] / ctl_out$sebetahat[1]

## I use this
multme <- solve(crossprod(X))[2, 2]
R <- qr.R(qr(X))
multme
(1 / R[2,2]) ^ 2

## Gagnon-Bartsch uses this
multgag <- solve(crossprod(cbind(X, vout$Zhat)))[2,2]

dat <- data_frame(SE1 = c(vout$sebetahat) * sqrt(multme / multgag),
                  SE2 = c(ctl_out$sebetahat))
ggplot(data = dat, mapping = aes(x = SE1, y = SE2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_bw() +
  ggtitle("Should lie on line")
```

Using the `ruv` R package:
```{r}
ruv4out <- ruv::RUV4(Y = Y, X = X[, 2, drop = FALSE], ctl = ctl, k = k,
                     Z = X[, -2, drop = FALSE])
ruv4_covadjust <- ruv::variance_adjust(ruv4out)

dat <- data_frame(GB = c(ruv4_covadjust$varbetahat.rsvar), ME = c(ctl_out$sebetahat) ^ 2)
ggplot(data = dat, mapping = aes(x = GB, y = ME)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_bw() +
  ggtitle("Should lie on line") +
  xlab("Gagnon-Barstch Var") +
  ylab("My Var")
```

# Summary
Let
\begin{align}
a := (X^TX)^{-1}[2,2]\\
b := ([X,Z]^T[X,Z])^{-1}[2,2]
\end{align}

When I calculate the variance inflation term during the maximum likelihood step, I end up using
\begin{align}
b\sigma_i^2 \frac{1}{m}\sum \frac{\beta_i^2}{a\sigma_i^2}\\
=\frac{b}{a}\sigma_i^2 \frac{1}{m}\sum \frac{\beta_i^2}{\sigma_i^2}.
\end{align}
whereas Gagnon-Bartsch uses
\begin{align}
b\sigma_i^2 \frac{1}{m}\sum \frac{\beta_i^2}{b\sigma_i^2}\\
=\sigma_i^2 \frac{1}{m}\sum \frac{\beta_i^2}{\sigma_i^2}.
\end{align}
Thus, it is slightly different.



```{r}
sessionInfo()
```
