---
title: "RUVB Plus ASH"
author: "David Gerard"
date: "January 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description
Here, I provide some comparisons between `seqgendiff` and the code from data_generators.R. I hope to show that they provide comparable comparisons.

I first take a look at only two_genes.

```{r}
## subset data -----------------------------------------------------------
## forces datamaker_counts and poisthin to have same effect sizes
set.seed(13)
constant_function <- function(n) { c(-1, 1) } 
library(seqgendiff)
mat <- read.csv("../output/gtex_tissue_gene_reads_v6p/muscle.csv")
max_row <- order(rowMeans(mat[, -(1:2)]), decreasing = TRUE)[1:2]
gmat <- mat[max_row, , drop = FALSE]
write.csv(gmat, "../output/small/muscle.csv", row.names = FALSE)
smalldat <- as.matrix(t(gmat[, -(1:2)]))

## settings for datamaker
source("../code/data_generators.R")
args <- list()
args$nullpi     <- 0
args$Nsamp      <- nrow(smalldat) / 2
args$Ngene      <- ncol(smalldat)
args$log2foldsd <- 1
args$poisthin   <- TRUE
args$tissue     <- "muscle"
args$path       <- "../output/small/"
args$alt_type   <- "constant" ## forces to have same effect size as poisthin


itermax <- 100
simmat <- matrix(NA, nrow = itermax, ncol = 4)
colnames(simmat) <- c("datamaker_1", "datamaker_neg1", "poisthin_neg1", "poisthin_1")
for (index in 1:itermax) {
  pout <- poisthin(mat = smalldat, nsamp = nrow(smalldat),
                 ngene = 2, gselect = "custom", gvec = c(TRUE, TRUE),
                 signal_fun = constant_function, signal_params = list(),
                 prop_null = 0)
  dout <- datamaker_counts_only(args)
  Yd <- log2(dout$input$counts + 1)
  Xd <- stats::model.matrix(~dout$input$condition)
  betad <- dout$meta$true_log2foldchange
  Yp <- log2(pout$Y + 1)
  Xp <- pout$X
  betap <- pout$beta
  lmd <- limma::lmFit(object = Yd, design = Xd)
  lmp <- limma::lmFit(object = t(Yp), design = Xp)
  simmat[index, 1:2] <- lmd$coefficients[, 2]
  simmat[index, 3:4] <- lmp$coefficients[, 2]
}
```

# Plot Results

Noteice that the boxplots should be centered about 1 and -1. datamaker is, and has much lower variability. Poisthin is scattered everywhere. This is a subtle bug and I need to figure out what's going on.
```{r}
boxplot(simmat)
abline(h = 1, lty = 2)
abline(h = -1, lty = 2)
```


Is it the way I am subsampling? I should observe half of the observations
```{r}
mean(Xp[, 2])
mean(Xd[, 2])
```

```{r}
sessionInfo()
```
