---
title: "Coverage Try 2"
author: "David Gerard"
date: "April 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

This is my second try at the coverage problem. Recall [here](ruvb_paper_plots_coverage.Rmd), I plotted the full boxplots of all of the coverage and calculated the median covarage. Here, I look at different loss functions.

# Read in data

```{r}
library(tidyverse)
library(stringr)
covdat <- read_csv(file = "../../reproduce_ruv3/Output/sims_out/cov_mat2.csv")
left_vals <- toupper(str_replace(str_extract(names(covdat)[-(1:5)], "^.+_"), "_", ""))
right_vals <- str_replace(str_extract(names(covdat)[-(1:5)], "_.+$"), "_", "")
name_vec <- c("Seed", "Pi0", "SampleSize", "NControls", "Poisthin",
              paste0(left_vals, right_vals))
name_vec[(length(name_vec) - 3):length(name_vec)] <- 
  paste0("RUVB", stringr::str_replace(string = names(covdat)[(length(name_vec) - 3):length(name_vec)],
                                      pattern = "(ruvb)(.*+)", replace = "\\2"))
names(covdat) <- name_vec
covdat <- select(covdat, -Seed, -Poisthin)
longdat <- gather(data = covdat, key = "Method", value = "Coverage", -(1:3))
```

Calculate losses on the data.

```{r}
less9 <- function(x) {
  s1 <- mean(x < 0.9)
}

g0975 <- function(x) {
  s2 <- mean(x > 0.975)
}

sumdat <- longdat %>% group_by(Pi0, SampleSize, NControls, Method) %>%
  summarise(Less = less9(Coverage), Greater = g0975(Coverage)) %>%
  ungroup() %>%
  mutate(Loss = Less + Greater)
```

Simple boxplot of results

```{r}
ggplot(data = sumdat, mapping = aes(x = Method, y = Loss)) +
  geom_boxplot() +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_rect(fill="white")) +
  ggtitle("Losses across all scenarios (sum < 0.9 and > 0.975)")

ggplot(data = sumdat, mapping = aes(x = Method, y = Less)) +
  geom_boxplot() +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_rect(fill="white")) +
  ggtitle("Proportion less than 0.9 across all scenarios")

ggplot(data = sumdat, mapping = aes(x = Method, y = Greater)) +
  geom_boxplot() +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_rect(fill="white")) +
  ggtitle("Proportion Greater than 0.975 across all scenarios")
```

Note that RUVB and RUVBnn have the best mean loss, but not the best mean Less(0.9) or the best mean Greater(0.975).
```{r}
meanlossdat <- sumdat %>% group_by(Method) %>%
  summarise(meanLoss = mean(Loss), meanLess = mean(Less), meanGreater = mean(Greater))
meanlossdat$Method[order(meanlossdat$meanLoss, decreasing = FALSE)]
meanlossdat$Method[order(meanlossdat$meanLess, decreasing = FALSE)]
meanlossdat$Method[order(meanlossdat$meanGreater, decreasing = FALSE)]
```

Try to combine the Less and Greater plots
```{r, fig.height=6, fig.width=9}
combdat <- select(sumdat, Pi0, SampleSize, NControls, Method, Less, Greater) %>%
  gather(key = "Loss", value = "Proportion", Less, Greater)

ggplot(data = combdat, mapping = aes(x = Method, y = Proportion, fill = Loss)) +
  geom_boxplot() + 
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(strip.background = element_rect(fill="white")) +
  ggtitle("Proportion less than 0.9 and Proportion greater than 0.975.")
```

Most of the methods perform significantly better on the "Less" criterion than the "Greater" criterion, but there are some methods where this is reveresed --- CATEo and CATEla. RUVB and RUVBnn perform about the same under both scenarios.

Connect dots for easier viewing. There is no method that has both a lower mean Greater loss and a lower mean Less loss.
```{r}
tempdat <- combdat %>% group_by(Method, Loss) %>%
  summarise(meanloss = mean(Proportion))
tempdat$RUVBorRUVBnn <- tempdat$Method == "RUVB" | tempdat$Method == "RUVBnn"

ggplot(data = tempdat, mapping = aes(x = Loss, y = meanloss, group = Method, color = RUVBorRUVBnn)) +
  geom_line() +
  theme_bw()
```

Repeat with by scenario rather than averaging over scenario.
```{r, fig.height=11, fig.width=9}
combdat$RUVBorRUVBnn <- combdat$Method == "RUVB" | combdat$Method == "RUVBnn"

factor_vec <- rep("other", length = nrow(combdat))
factor_vec[stringr::str_detect(combdat$Method, "c$")] <- "c"
factor_vec[stringr::str_detect(combdat$Method, "m$")] <- "m"
factor_vec[combdat$Method == "RUVB" | combdat$Method == "RUVBnn"] <- "RUVB"
combdat$categories <- as.factor(factor_vec)

ggplot(data = combdat, mapping = aes(x = Loss, y = Proportion, group = Method, color = RUVBorRUVBnn)) +
  geom_line() +
  facet_grid(Pi0 + NControls ~ SampleSize) +
  theme_bw() +
  theme(strip.background = element_rect(fill="white")) 

ggplot(data = combdat, mapping = aes(x = Loss, y = Proportion, group = Method, color = factor_vec)) +
  geom_line() +
  facet_grid(Pi0 + NControls ~ SampleSize) +
  theme_bw() +
  theme(strip.background = element_rect(fill="white")) 
```


# Summary

* RUVB and RUVBnn perform best under large sample sizes when there are few control genes. But it seems to perform about the same under every scenario. 
* RUVB and RUVBnn have equal losses for the less than 0.9 and the greater than 0.975 criteria. No other method has equal losses here.
* Using this criteria, other methods do seem to have better coverage for many of the scenarios, but they are not the same methods for all scenarios (see boxplots).
* Using this criteron, RUVB seems to be the best method when there are few control genes and the proportion of null genes is 0.5.
* "m" methods are very very conservative when Pi0 = 0.5 and are very anti-conservative when pi0 =  1. They perform very very well when Pi0 = 0.9.
* "c" methods perform very pooorly when m = 10 or when n is small. They tend to be too conservative.

```{r}
sessionInfo()
```
