---
title: "Differences in `sprop` and `var_inflate_pen`"
author: "David Gerard"
date: "May 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I look at differences in using the `sprop` and `var_inflate_pen` parameters in `vicar::mouthwash`.

# Analysis

First, we obtain simulated data with some signal. The original Gtex data were processed in the same pipeline as [here](https://github.com/dcgerard/ruvb_sims).

```{r}
set.seed(9)
ngene <- 1000
nsamp <- 20
nullpi <- 0.5
log2foldsd <- 3

mat <- t(as.matrix(read.csv("../../reproduce_ruv3/Output/gtex_tissue_gene_reads_v6p/muscle.csv",
                            header = TRUE)[, -c(1,2)]))
submat <- mat[, order(apply(mat, 2, median), decreasing = TRUE)[1:ngene]]
rm(mat)
d_out <- seqgendiff::poisthin(mat = submat,
                              nsamp = nsamp,
                              ngene = ngene,
                              signal_params = list(mean = 0, sd = log2foldsd),
                              gvec = rep(TRUE, length(ngene)),
                              gselect = "custom",
                              prop_null = nullpi,
                              alpha = 1)
which_null <- abs(d_out$beta) < 10 ^ -6
nnull         <- sum(which_null)

beta_true <- d_out$beta
X <- d_out$X
colnames(X) <- c("Intercept", "Treatment")
Y <- log2(d_out$Y + 1)
num_sv <- max(sva::num.sv(t(Y), mod = X, method = "be"), 1)
```

Now, we fit mouthwash under different levels of `sprop` and `var_inflate_pen`.

```{r}
library(vicar)
mout_0_0 <- mouthwash(Y = Y, X = X, k = num_sv, cov_of_interest = 2,
                      include_intercept = FALSE, sprop = 0, var_inflate_pen = 0)
mout_0_1 <- mouthwash(Y = Y, X = X, k = num_sv, cov_of_interest = 2,
                      include_intercept = FALSE, sprop = 0, var_inflate_pen = 1)
mout_1_1 <- mouthwash(Y = Y, X = X, k = num_sv, cov_of_interest = 2,
                      include_intercept = FALSE, sprop = 1, var_inflate_pen = 1)
mout_1_0001 <- mouthwash(Y = Y, X = X, k = num_sv, cov_of_interest = 2,
                      include_intercept = FALSE, sprop = 1, var_inflate_pen = 0.0001)
```

# Compare fits

First, there is more similarity in the variability in estimates of `xi` between different `sprop`'s than between different `var_inflate_pen`'s.
```{r}
mout_0_0$xi 
mout_0_1$xi 
mout_1_0001$xi 
mout_1_1$xi 
```

Second, when we use the correct level of `sprop`, we get much better estimates of `pi0`.

```{r}
mout_0_0$pi0
mout_0_1$pi0
mout_1_0001$pi0
mout_1_1$pi0
```

These are results for just one dataset, but they seem to occur in general.

```{r}
sessionInfo()
```
