---
title: "Compare MOUTHWASH Subsampling"
author: "David Gerard"
date: "May 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I look at differences in using `subsample = TRUE` vs `subsample = FALSE` in `vicar::mouthwash`.

# Simulate Data

```{r}
library(tidyverse)
library(seqgendiff)
library(vicar)
mat <- t(read.csv("../../reproduce_ruv3/Output/gtex_tissue_gene_reads_v6p/blood.csv")[, -c(1,2)])
dout <- seqgendiff::poisthin(mat = mat, nsamp = 30, ngene = 10000, prop_null = 0.5, 
                             signal_fun = stats::rnorm, signal_params = list(mean = 0, sd = 1))
Y <- log2(dout$Y + 1)
X <- dout$X
```

# Fit both versions of mouthwash and compare

Estimate number of hidden confounders.
```{r}
num_sv <- sva::num.sv(dat = t(Y), mod = X)
num_sv
```

Full mouthwash.
```{r, cache=TRUE}
fstart <- proc.time()
mout_full <- mouthwash(Y = Y, X = X, k = num_sv, 
                       cov_of_interest = 2, include_intercept = FALSE)
ftot <- proc.time() - fstart
```

Subsample mouthwash.
```{r, cache=TRUE}
sstart <- proc.time()
mout_sub <- mouthwash(Y = Y, X = X, k = num_sv, 
                      cov_of_interest = 2, include_intercept = FALSE, 
                      subsample = TRUE, num_sub = 1000)
stot <- proc.time() - sstart
```

Fit `limma::voom` -> `limma::limma` -> `limma:ebayes` -> `ashr::ash` pipeline for good measure:
```{r, cache=TRUE}
astart     <- proc.time()
voom_out   <- limma::voom(counts = t(dout$Y), design = X)
limma_out  <- limma::lmFit(object = voom_out)
ebayes_out <- limma::ebayes(fit = limma_out)
aout       <- ashr::ash.workhorse(betahat = limma_out$coefficients[, 2],
                                  sebetahat = sqrt(ebayes_out$s2.post),
                                  df = ebayes_out$df.total)
atot       <- proc.time() - astart
```

# Compare methods

Runtimes
```{r}
temp <- matrix(c(ftot[1], stot[1], atot[1]), ncol = 3)
colnames(temp) <- c("Full MOUTHWASH", "Sub MOUTHWASH", "ASH")
knitr::kable(temp)
```

Subsample mouthwash gets the estimate of pi0 pretty good!
```{r}
temp <- matrix(c(ashr::get_pi0(mout_full),
                 ashr::get_pi0(mout_sub),
                 ashr::get_pi0(aout)),
               ncol = 3)
colnames(temp) <- c("Full MOUTHWASH", "Sub MOUTHWASH", "ASH")
knitr::kable(temp)
```

Sub mouthwash and full mouthwash have almost the exact same posterior means
```{r}
dat <- data_frame(full = mout_full$result$PosteriorMean, sub = mout_sub$result$PosteriorMean)
ggplot(data = dat, mapping = aes(x = full, y = sub)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "blue", lty = 2, lwd = 1) +
  theme_bw() +
  ggtitle("Posterior Means")
```

lfdr's are a little bit more different
```{r}
dat <- data_frame(full = mout_full$result$lfdr, sub = mout_sub$result$lfdr)
ggplot(data = dat, mapping = aes(x = full, y = sub)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "blue", lty = 2, lwd = 1) +
  theme_bw() +
  ggtitle("lfdr's")
```

The different estimate of the hidden confounders:
```{r}
mout_full$z2
mout_sub$z2
```

But there is only a small decrease in AUC:
```{r}
library(pROC)
which_null <- abs(dout$beta) < 10 ^ -10
aucmat <- matrix(c(pROC::auc(pROC::roc(predictor = mout_full$result$lfdr, response = which_null)),
                   pROC::auc(pROC::roc(predictor = mout_sub$result$lfdr, response = which_null)),
                   pROC::auc(pROC::roc(predictor = aout$result$lfdr, response = which_null))),
                 ncol = 3)
colnames(aucmat) <- c("Full MOUTHWASH", "Sub MOUTHWASH", "ASH")
knitr::kable(aucmat)
```


```{r}
sessionInfo()
```
