---
title: "Improved MOUTHWASH Simulations Analysis"
author: "David Gerard"
date: "May 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I run the mouthwash simulations under $\alpha = 0$. Most methods are now the limma-shrunk methods. I additionally have:

1. Made it so that CATErr uses the truncated SVD as its factor analysis.
2. Included the normal-likelihood version of MOUTHWASH with a uniform grid.

The simulations were implemented [here](https://github.com/dcgerard/ruvb_plus_ash/blob/master/code/mouthwash_sims_alphais0_longer.R) with code [here](https://github.com/dcgerard/ruvb_plus_ash/blob/master/code/non_nc_methods.R) and [here](https://github.com/dcgerard/ruvb_sims/blob/nc_var/Code/nc_adjustment_methods.R).

# Summary

1. There is a uniform grid effect. Using a uniform grid makes the normal likelihood more conservative. This might indicate that I need to increase my grid size.
2. The $t$ is just a little more conservative. I wouldn't necessarily call this "better" or "worse".
3. In terms of AUC, the $t$ performs as well as the normal in terms of median auc. In terms of mean auc, it performs a little worse.
4. Though, the normal likelihood with a uniform grid appears to perform just as well as the normal likelihood with a normal grid. Since the only difference between the normal and t with a uniform grid is the former has `df = Inf`, I think this is evidence that my code is correct.

# $\pi_0$
```{r, fig.height=9, fig.width=9}
library(tidyverse)
dat <- as_data_frame(readRDS(file = "../output/alpha_1_sims_out/sims_out_alpha0_moreruns.RDS"))

subdat <- select(dat, nullpi, Nsamp, contains("pi0_")) %>%
  gather(key = "Method", value = "pi0hat", contains("pi0"))
subdat$Method <- stringr::str_replace(subdat$Method, "pi0_", "")
subdat$type <- rep("other", nrow(subdat))
subdat$type[stringr::str_detect(subdat$Method, "m$")] <- "MAD"
subdat$type[stringr::str_detect(subdat$Method, "norm$")] <- "other"
subdat$type[stringr::str_detect(subdat$Method, "c$")] <- "ctl"

ggplot(data = subdat, mapping = aes(y = pi0hat, x = Method, fill = type)) +
  facet_grid(Nsamp~nullpi) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.background = element_rect(fill = "white")) +
  geom_hline(mapping = aes(yintercept = nullpi), lty = 2, lwd = 1, color = "gray50") +
  ggtitle("Pi0hat")
```

From the above plot, MOUTHWASH methods do the best in terms of calibration.

Let's look at MOUTHWASH methods in more detail. We have the following key:

- norm = normal mouthwash with normal mixture prior fit with $\alpha = 0$
- norm_unif = normal mouthwash with uniform mixture prior fit with $\alpha = 0$
- t = t mouthwash with uniform mixture prior fit with $\alpha = 0$

```{r}
mouth_dat <- filter(subdat, stringr::str_detect(Method, "mouthwash"))
mouth_dat$Method <- stringr::str_replace(mouth_dat$Method, "ash_mouthwash_", "")

ggplot(data = mouth_dat, mapping = aes(y = pi0hat, x = Method)) +
  facet_grid(Nsamp~nullpi) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.background = element_rect(fill = "white")) +
  geom_hline(mapping = aes(yintercept = nullpi), lty = 2, lwd = 1, color = "gray50") +
  ggtitle("Pi0hat for MOUTHWASH")

```

Thoughts:

1. The $t$ is just a little more conservative.
2. There is definitely a uniform grid effect. The normal with a uniform grid lies between a t with a uniform grid and a normal with a normal grid.


# AUC

```{r, fig.height=9, fig.width=9}
rm(list = ls())
dat <- as_data_frame(readRDS(file = "../output/alpha_1_sims_out/sims_out_alpha0_moreruns.RDS"))

subdat <- select(dat, nullpi, Nsamp, contains("auc_")) %>%
  gather(key = "Method", value = "auc", contains("auc"))
subdat$Method <- stringr::str_replace(subdat$Method, "auc_", "")
subdat$type <- rep("other", nrow(subdat))
subdat$type[stringr::str_detect(subdat$Method, "m$")] <- "MAD"
subdat$type[stringr::str_detect(subdat$Method, "norm$")] <- "other"
subdat$type[stringr::str_detect(subdat$Method, "c$")] <- "ctl"

ggplot(data = subdat, mapping = aes(y = auc, x = Method, fill = type)) +
  facet_grid(Nsamp~nullpi) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.background = element_rect(fill = "white")) +
  ggtitle("auc")
```

```{r, fig.height=9, fig.width=9}
sumdat <- subdat %>% group_by(Method, Nsamp, nullpi) %>%
  summarize(med_auc = median(auc)) %>%
  ungroup()
sumdat$type <- rep("other", nrow(sumdat))
sumdat$type[stringr::str_detect(sumdat$Method, "m$")] <- "MAD"
sumdat$type[stringr::str_detect(sumdat$Method, "norm$")] <- "other"
sumdat$type[stringr::str_detect(sumdat$Method, "c$")] <- "ctl"

maxdat <- sumdat %>% group_by(Nsamp, nullpi) %>%
  summarize(max_auc = max(med_auc)) %>%
  ungroup()

ggplot(data = sumdat, mapping = aes(x = Method, y = med_auc, color = type)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.background = element_rect(fill = "white")) +
  facet_grid(Nsamp ~ nullpi) +
  geom_hline(data = maxdat, mapping = aes(yintercept = max_auc), lty = 2) +
  ylab("Median AUC") +
  ggtitle("Median AUC")
```


Look at just the mouthwash methods.

1. The $t$ performs only slightly worse than the normal when $n = 6$ ad about equivalently for $n = 10$ or $40$.
2. The normal uniform performs just as well as the normal normal at $n = 6$.


```{r}
mouth_dat <- filter(sumdat, stringr::str_detect(Method, "mouthwash"))
mouth_dat$Method <- stringr::str_replace(mouth_dat$Method, "ash_mouthwash_", "")

ggplot(data = mouth_dat, mapping = aes(x = Method, y = med_auc)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.background = element_rect(fill = "white")) +
  facet_grid(Nsamp ~ nullpi, scales = "free_y") +
  geom_hline(data = maxdat, mapping = aes(yintercept = max_auc), lty = 2) +
  ylab("Median AUC") +
  ggtitle("Median AUC")
```

We'll calculate bootstrap intervals real quick

```{r}
mouth_sub <- filter(subdat, stringr::str_detect(Method, "mouthwash"))

boot_med <- function(x, boot_num = 1000) {
  bvec <- rep(NA, length = boot_num)
  for (bindex in 1:boot_num) {
    bvec[bindex] <- median(sample(x, replace = TRUE))
  }
  return(c(median(x), quantile(bvec, probs = c(0.025, 0.975))))
}

boot_out <- group_by(mouth_sub, Method, nullpi, Nsamp) %>%
  do(setNames(data.frame(t(boot_med(.$auc))), c("median", "lower", "upper")))

boot_out$Method <- stringr::str_replace(boot_out$Method, "ash_mouthwash_", "")
```


The CI's definitely still overlap. It seems that the normal and t now perform about equally. But this is in terms of *median* auc. The next section recapitulates my results with *mean* auc.

```{r}
ggplot(data = boot_out, mapping = aes(y = median, x = Method, ymin = lower, ymax = upper)) +
  geom_point() +
  geom_linerange() +
  theme_bw() +
  facet_grid(Nsamp ~ nullpi, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), 
        strip.background = element_rect(fill = "white")) +
  geom_hline(data = maxdat, mapping = aes(yintercept = max_auc), lty = 2)
```


# Mean AUC

```{r, fig.height=9, fig.width=9}
rm(list = ls())
dat <- as_data_frame(readRDS(file = "../output/alpha_1_sims_out/sims_out_alpha0_moreruns.RDS"))

subdat <- select(dat, nullpi, Nsamp, contains("auc_")) %>%
  gather(key = "Method", value = "auc", contains("auc"))
subdat$Method <- stringr::str_replace(subdat$Method, "auc_", "")
subdat$type <- rep("other", nrow(subdat))
subdat$type[stringr::str_detect(subdat$Method, "m$")] <- "MAD"
subdat$type[stringr::str_detect(subdat$Method, "norm$")] <- "other"
subdat$type[stringr::str_detect(subdat$Method, "c$")] <- "ctl"

sumdat <- subdat %>% group_by(Method, Nsamp, nullpi) %>%
  summarize(med_auc = mean(auc), sd_auc = sd(auc), n = length(auc)) %>%
  ungroup() %>%
  mutate(lower = med_auc - 1.96 * sd_auc / sqrt(n),
         upper = med_auc + 1.96 * sd_auc / sqrt(n))
sumdat$type <- rep("other", nrow(sumdat))
sumdat$type[stringr::str_detect(sumdat$Method, "m$")] <- "MAD"
sumdat$type[stringr::str_detect(sumdat$Method, "norm$")] <- "other"
sumdat$type[stringr::str_detect(sumdat$Method, "c$")] <- "ctl"

maxdat <- sumdat %>% group_by(Nsamp, nullpi) %>%
  summarize(max_auc = max(med_auc)) %>%
  ungroup()

ggplot(data = sumdat, mapping = aes(x = Method, y = med_auc, color = type)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.background = element_rect(fill = "white")) +
  facet_grid(Nsamp ~ nullpi) +
  geom_hline(data = maxdat, mapping = aes(yintercept = max_auc), lty = 2) +
  ylab("mean AUC") +
  ggtitle("mean AUC")
```


Look at just the mouthwash methods.

1. The $t$ performs a little worse than the "normal" in terms of mean AUC. And this appears to be statistically significant. But the normal with a uniform grid performs just as well. So there doesn't seem to be a coding error here.


```{r}
mouth_dat <- filter(sumdat, stringr::str_detect(Method, "mouthwash"))
mouth_dat$Method <- stringr::str_replace(mouth_dat$Method, "ash_mouthwash_", "")

ggplot(data = mouth_dat, mapping = aes(x = Method, y = med_auc)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.background = element_rect(fill = "white")) +
  facet_grid(Nsamp ~ nullpi, scales = "free_y") +
  geom_hline(data = maxdat, mapping = aes(yintercept = max_auc), lty = 2) +
  ylab("mean AUC") +
  ggtitle("mean AUC") +
  geom_linerange(mapping = aes(ymin = lower, ymax = upper))
```
```{r}
sessionInfo()
```
