---
title: "RUVB vs All Loss Plots"
author: "David Gerard"
date: "`r Sys.Date()`"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

# Abstract

Here, I make a bunch of lossplots of RUVB vs everything else.

# Summary
After looking at the lossplots, here are my thoughts:

* RUVB and RUVBnn have almost the exact same coverage.
* RUV2o and RUV3o work really well and outperform RUVB for small n.
* RUV3la and RUV2l are competitive but no where completely beat RUVB.
* No other method is competitive everywhere.

# Key

Here, I explore six mean methods:

* OLS (no confounder adjustment)
* CATE
* RUV4
* RUV3
* RUV2
* RUVB

I estimate the variances with the mean methods in different combinations. The key for this is:

* o = Original.
* n = $t$ approximation for the likelihood in RUVB.
* nn = normal approximation for the likelihood in RUVB.
* m = MAD variance inflation.
* c = control-gene calibrated.
* l = Limma-adjusted.
* lb = limma-adjusted Before GLS (for either CATE or RUV3).
* la = limma-adjusted After GLS (for either CATE or RUV3).
* d = Delta-adjustment from CATE package (additive variance inflation).

# Boxplots

## Read in Data

```{r, message=FALSE}
library(tidyverse)
library(stringr)
covdat <- read_csv(file = "../../reproduce_ruv3/Output/sims_out/cov_mat2.csv")
left_vals <- toupper(str_replace(str_extract(names(covdat)[-(1:5)], "^.+_"), "_", ""))
right_vals <- str_replace(str_extract(names(covdat)[-(1:5)], "_.+$"), "_", "")
name_vec <- c("Seed", "Pi0", "SampleSize", "NControls", "Poisthin",
              paste0(left_vals, right_vals))
name_vec[(length(name_vec) - 3):length(name_vec)] <-
  paste0("RUVB", stringr::str_replace(string = names(covdat)[(length(name_vec) - 3):length(name_vec)],
                                      pattern = "(ruvb)(.*+)", replace = "\\2"))
names(covdat) <- name_vec
covdat <- select(covdat, -Seed, -Poisthin)

other_methods <- names(covdat)[-(1:3)]
other_methods <- other_methods[!stringr::str_detect(other_methods, "RUVB")]

which_ruvb <- stringr::str_detect(names(covdat), "RUVB")
```

## Now the lossplots:

```{r, fig.height=11, fig.width=9}
less9 <- function(x) {
  s1 <- mean(x < 0.9)
}

g0975 <- function(x) {
  s2 <- mean(x > 0.975)
}

for (index in 1:length(other_methods)) {
  which_columns <- names(covdat) == other_methods[index] | which_ruvb
  which_columns[1:3] <- TRUE
  longdat <- gather(data = covdat[which_columns], key = "Method", value = "Coverage", -(1:3))
  sumdat <- longdat %>% group_by(Pi0, SampleSize, NControls, Method) %>%
    summarise(Less = less9(Coverage), Greater = g0975(Coverage)) %>%
    ungroup()
  combdat <- select(sumdat, Pi0, SampleSize, NControls, Method, Less, Greater) %>%
    gather(key = "Loss", value = "Proportion", Less, Greater)
  
  pl <- ggplot(data = combdat, mapping = aes(x = Loss, y = Proportion,
                                             group = Method, color = Method)) +
    geom_line() +
    facet_grid(Pi0 + NControls ~ SampleSize) +
    theme_bw() +
    theme(strip.background = element_rect(fill="white")) +
    coord_cartesian(xlim = c(1.45, 1.55)) +
    xlab("Loss Type") +
    ylab("Loss") +
    ggtitle(other_methods[index])
  print(pl)
}
```

# Session information

<!-- Insert the session information into the document -->
```{r session-info}
sessionInfo()
```
