---
title: "RUVB LFSR's"
author: "David Gerard"
date: "January 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Abstract

This is meant just to look at a few datasets to see if RUVB tentatively calculates good lfsr's. RUVB doesn't return good lfsr's but MOUTHWASH does. Running ASH on RUVB is competitive with MOUTHWASH.

# Generate Data and Fit Method

Generate Poisson-thinned data under the simulation settings used in paper.

```{r}
source("../code/data_generators.R")
args <- list()
args$nullpi       <- 0.9
args$Nsamp        <- 10
args$ncontrol     <- 100
args$poisthin     <- TRUE
args$log2foldsd   <- 0.8
args$tissue       <- "muscle"
args$path         <- "../output/gtex_tissue_gene_reads_v6p/"
args$Ngene        <- 1000
args$skip_gene    <- 0
args$log2foldmean <- 0

set.seed(723)
d_out <- datamaker_counts_only(args)
which_null <- d_out$meta$null
control_genes <- as.logical(which_null)
nnull         <- sum(control_genes)
control_genes[control_genes][sample(1:nnull, size = nnull - args$ncontrol)] <- FALSE
beta_true <- rep(0, length = args$Ngene)
beta_true[!which_null] <- d_out$meta$true_log2foldchange
X <- as.matrix(model.matrix(~d_out$input$condition))
colnames(X) <- c("Intercept", "Treatment")
Y <- t(log2(as.matrix(d_out$input$counts + 1)))
num_sv <- max(sva::num.sv(t(Y), mod = X, method = "be"), 1)
```

Fit RUVB

```{r}
library(vicar)
library(ggplot2)
ruvbout <- ruvb(Y = Y, X = X, ctl = control_genes, k = num_sv,
                fa_args = list(use_code = "r", nsamp = 10000, display_progress = FALSE),
                cov_of_interest = 2)

mout <- mouthwash(Y = Y, X = X, k = num_sv, cov_of_interest = 2)

morder <- order(mout$result$qvalue)
m_fdr <- cumsum(which_null[morder]) / (1:length(which_null))
m_qvalue_sorted <- mout$result$qvalue[morder]
mdat <- data.frame(m_fdr, m_qvalue = m_qvalue_sorted)
ggplot(data = mdat, mapping = aes(x = m_qvalue, y = m_fdr)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("MOUTHWASH")


rorder <- order(ruvbout$lfsr2, na.last = NA)
r_fdr <- cumsum(which_null[rorder]) / (1:sum(!is.na(ruvbout$lfsr2)))
r_svalue_sorted <- ruvbout$svalues2[rorder]
rdat <- data.frame(r_fdr, r_svalue = r_svalue_sorted)
ggplot(data = rdat, mapping = aes(x = r_svalue, y = r_fdr)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("RUVB")
```

MOUTHWASH does awesome, RUVB does not do well. Now run ASH on RUVB.

```{r}
aout <- ashr::ash(betahat = c(ruvbout$means), sebetahat = c(ruvbout$sd))
alfdr <- ashr::get_lfdr(aout)
aorder <- order(alfdr)
a_fdr <- cumsum(which_null[aorder]) / 1:length(aorder)
a_qvalue <- ashr::get_qvalue(aout)[aorder]
adat <- data.frame(a_fdr, a_qvalue)
ggplot(data = adat, mapping = aes(x = a_qvalue, y = a_fdr)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("RUVB + ASH")
```


RUVB + ASH does really well.

```{r}
sessionInfo()
```
