---
title: "RUVB Plus ASH"
author: "David Gerard"
date: "January 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis

Read in estimates of $\pi_0$ and calculate mean of estimates of $\pi_0$.

```{r read}
library(dplyr)
pi0hat <- read.csv(file = "../output/sims_out/pi0_mat2.csv")
pos_bad <- grep(names(pi0hat), pattern = "auc") # some columns accidently have auc instead of pi0hat
pi0hat <- select(pi0hat, -pos_bad) %>% select(-pi0_qvalue_ruvb)
mean_dat <- pi0hat %>%
  dplyr::select(-current_seed, -poisthin) %>%
  dplyr::group_by(nullpi, Nsamp, ncontrols) %>%
  dplyr::summarise_each(funs(mean))
names(mean_dat) <- stringr::str_replace(names(mean_dat), pattern = "pi0_", replacement = "")
```

Order columns by mse from truth of 0.9.

```{r}
colorder <- order(apply(mean_dat[, -(1:3)], 2, function(x) {sum((x - 0.9)^2)}), decreasing = FALSE)
method_order <- names(mean_dat)[-(1:3)][colorder]
colorder <- c(1:3, colorder + 3)
data.table::setcolorder(mean_dat, colorder)
```

Data table of results.

```{r}
mean_dat_round <- mean_dat
mean_dat_round[, -(1:3)] <- round(mean_dat_round[, -(1:3)], digits = 2)
DT::datatable(mean_dat_round, extensions = 'FixedColumns',
              options = list(dom = "t", scrollX = TRUE, fixedColumns = list(leftColumns = 3)),
              rownames = FALSE)
```

## Plot estimates

RUVB doesn't do as well as MOUTHWASH and BACKWASH. But in terms of ASH methods, the only other one that does better than RUVB is RUV3.

```{r plot_pi0, fig.height=15}
library(ggplot2)
library(tidyr)
longdat <- gather(mean_dat, key = method, value = pi0hat, 4:ncol(mean_dat))
longdat$ncontrols <- as.factor(longdat$ncontrols)
longdat$method <- factor(longdat$method, levels = method_order)
ggplot(data = longdat, mapping = aes(y = pi0hat, x = Nsamp, color = ncontrols)) +
  facet_wrap(~method, ncol = 3) +
  geom_line() + 
  geom_hline(yintercept = 0.9, lty = 2) +
  theme_bw() + 
  ggtitle("Mean Estimates of Pi0")
```


## Same thing but with AUC

```{r}
rm(list = ls())
auchat <- read.csv(file = "../output/sims_out/auc_mat2.csv")
pos_bad <- grep(names(auchat), pattern = "mse") # some columns accidently have mse instead of auc
auchat <- select(auchat, -pos_bad)
mean_dat <- auchat %>%
  dplyr::select(-current_seed, -poisthin) %>%
  dplyr::group_by(nullpi, Nsamp, ncontrols) %>%
  dplyr::summarise_each(funs(mean))
names(mean_dat) <- stringr::str_replace(names(mean_dat), pattern = "auc_", replacement = "")

```

Order columns by mse from 1.

```{r}
colorder <- order(apply(mean_dat[, -(1:3)], 2, function(x) {sum((x - 1)^2)}), decreasing = FALSE)
method_order <- names(mean_dat)[-(1:3)][colorder]
colorder <- c(1:3, colorder + 3)
data.table::setcolorder(mean_dat, colorder)
```

Data table of results.

```{r}
mean_dat_round <- mean_dat
mean_dat_round[, -(1:3)] <- round(mean_dat_round[, -(1:3)], digits = 2)
DT::datatable(mean_dat_round, extensions = 'FixedColumns',
              options = list(dom = "t", scrollX = TRUE, fixedColumns = list(leftColumns = 3)),
              rownames = FALSE)
```

RUVB seems to do 4th or 5th best behind both mouthwash methods, backwash, and limma-shrunk/variance inflated ruv4. Interestingly, its AUC seems to be rather robust to the number of control genes used, unlike other control gene methods.

```{r plot_auc, fig.height=15}
library(ggplot2)
library(tidyr)
longdat <- gather(mean_dat, key = method, value = pi0hat, 4:ncol(mean_dat))
longdat$ncontrols <- as.factor(longdat$ncontrols)
longdat$method <- factor(longdat$method, levels = method_order)
ggplot(data = longdat, mapping = aes(y = pi0hat, x = Nsamp, color = ncontrols)) +
  facet_wrap(~method, ncol = 3) +
  geom_line() + 
  geom_hline(yintercept = 1, lty = 2) +
  theme_bw() +
  ggtitle("Mean AUC")
```

```{r}
sessionInfo()
```
