---
title: "RUVB vs All Boxplots"
author: "David Gerard"
date: "`r Sys.Date()`"
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

# Abstract

Here, I make a bunch of boxplots of coverage with RUVB and just one other method. These are just a bunch of plots so that I can look at them.

# Summary
After looking at the boxplots, these are my thoughts:

* As I noted before, MAD adjustments (green) are conservative for Pi0 = 0.5 and anticonservative for Pi0 = 1. This is true for all methods.
* Control gene variance adjustments (red) do not work when there are few control genes.
* None of the RUV4-type methods perform unbiasedly in all scenarios (though the CATEd comes close with only being conservative when n is large and there are few control genes).
* I will now list the methods that just by eye seem like they perform competitively. By this I mean they don't seem too conservative or anticonservative in any scenario. They are RUVB, RUVBnn, RUV2, RUV2l, RUV3o, RUV3la (NOT RUV3lb).
* The best RUVB methods are RUVB and RUVBnn (not the t-based methods).

# Key

Here, I explore six mean methods:

* OLS (no confounder adjustment)
* CATE
* RUV4
* RUV3
* RUV2
* RUVB

I estimate the variances with the mean methods in different combinations. The key for this is:

* o = Original.
* n = $t$ approximation for the likelihood in RUVB.
* nn = normal approximation for the likelihood in RUVB.
* m = MAD variance inflation.
* c = control-gene calibrated.
* l = Limma-adjusted.
* lb = limma-adjusted Before GLS (for either CATE or RUV3).
* la = limma-adjusted After GLS (for either CATE or RUV3).
* d = Delta-adjustment from CATE package (additive variance inflation).

# Boxplots

## Read in Data

```{r, message=FALSE}
library(tidyverse)
library(stringr)
covdat <- read_csv(file = "../../reproduce_ruv3/Output/sims_out/cov_mat2.csv")
left_vals <- toupper(str_replace(str_extract(names(covdat)[-(1:5)], "^.+_"), "_", ""))
right_vals <- str_replace(str_extract(names(covdat)[-(1:5)], "_.+$"), "_", "")
name_vec <- c("Seed", "Pi0", "SampleSize", "NControls", "Poisthin",
              paste0(left_vals, right_vals))
name_vec[(length(name_vec) - 3):length(name_vec)] <-
  paste0("RUVB", stringr::str_replace(string = names(covdat)[(length(name_vec) - 3):length(name_vec)],
                                      pattern = "(ruvb)(.*+)", replace = "\\2"))
names(covdat) <- name_vec
covdat <- select(covdat, -Seed, -Poisthin)

other_methods <- names(covdat)[-(1:3)]
other_methods <- other_methods[!stringr::str_detect(other_methods, "RUVB")]

which_ruvb <- stringr::str_detect(names(covdat), "RUVB")
```

## Now the boxplots:

```{r, fig.height=11, fig.width=9}
for (index in 1:length(other_methods)) {
  which_columns <- names(covdat) == other_methods[index] | which_ruvb
  which_columns[1:3] <- TRUE
  longdat <- gather(data = covdat[which_columns], key = "Method", value = "Coverage", -(1:3))
  
  if (stringr::str_detect(other_methods[index], "m")) {
    color_vec <- c("green", "purple")
  } else if (stringr::str_detect(other_methods[index], "c")) {
    color_vec <- c("red", "purple")
  } else {
    color_vec <- c("blue", "purple")
  }
  names(color_vec) <- c(other_methods[index], "RUVB")  
  log_vec <- stringr::str_detect(longdat$Method, "RUVB")
  category <- rep("RUVB", length = nrow(longdat))
  category[!log_vec] <- other_methods[index]
  longdat$category <- category
  
  pl <- ggplot(data = longdat, mapping = aes(y = Coverage, x = Method, fill = category)) +
    geom_boxplot(outlier.size = 0.2, size = 0.2) +
    facet_grid(Pi0 + NControls ~ SampleSize) +
    geom_hline(yintercept = 0.95, lty = 2) +
    theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(strip.background = element_rect(fill="white")) +
    ggtitle(paste0(other_methods[index])) +
    scale_fill_manual(values = color_vec)
  print(pl)
}
```

# Session information

<!-- Insert the session information into the document -->
```{r session-info}
sessionInfo()
```
