---
title: "Variance Estimation Combinations"
author: "David Gerard"
date: "April 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

This write-up is meant for me to code up all possible combinations of variance estimation and confounder adjustment approaches.

# Simulate Data

```{r}
set.seed(545)
n <- 11 ## number of samples
p <- 101 ## number of genes
q <- 3 ## number of confounders
k <- 2 ## number of observed covariates
m <- 23 ## number of controls

sd_vec <- sqrt(rchisq(p, df = 5) / 5)

is_zero <- rep(FALSE, length = p)
is_zero[1:71] <- TRUE ## about 70% of genes are null
ctl <- rep(FALSE, length = p)
ctl[1:m] <- TRUE

X <- matrix(rnorm(n * k), nrow = n)
B <- matrix(rnorm(k * p), ncol = p)
B[2, is_zero] <- 0
Z <- matrix(rnorm(n * q), nrow = n)
A <- matrix(rnorm(q * p), ncol = p)
E <- matrix(rnorm(n * p), nrow = n) %*% diag(sd_vec)

Y <- X %*% B + Z %*% A + E
```

# RUV built-in functions.

The ruv R package provides a few built-in methods for estimating the variances.

```{r}
ruv2out <- ruv::RUV2(Y = Y, X = X[, 2, drop = FALSE],
                     ctl = ctl, k = q, Z = X[, 1, drop = FALSE])
ruv4out <- ruv::RUV4(Y = Y, X = X[, 2, drop = FALSE],
                     ctl = ctl, k = q, Z = X[, 1, drop = FALSE])

ruv2varout <- ruv::variance_adjust(fit = ruv2out)
ruv4varout <- ruv::variance_adjust(fit = ruv4out)
```

Look and see the difference in the different estimates
```{r}
hist(ruv2varout$varbetahat) ## The default estimate
hist(ruv2varout$varbetahat.rsvar) ## rescaled estimate using control genes
hist(ruv2varout$varbetahat.ebayes) ## limma-shrunk variances
hist(ruv2varout$varbetahat.rsvar.ebayes) ## rescaled estimate using control genes of limma-shrunk variances
```

Demonstrate that `ruv2varout$varbetahat.ebayes` is just limmashrunk form OLS variance estimates.
```{r}
limma_adjusted <- limma::squeezeVar(var = ruv2varout$varbetahat, df = ruv2varout$df)
max(limma_adjusted$var.post - c(ruv2varout$varbetahat.ebayes))
```

Re-derive variance scaling by hand

```{r}
rsvar_hand <- mean(ruv2varout$betahat[ctl] ^ 2 / ruv2varout$sigma2[ctl]) * ruv2varout$sigma2
max(rsvar_hand - ruv2varout$varbetahat.rsvar)
```

# All of the different CATE

You can change the parameter `nc.var.correction` to be `TRUE` or `FALSE` for CATE's additive correction.

You can also set `calibrate` to be `TRUE` or `FALSE` for median centering and MAD inflation.

```{r}
cout_pre <- cate::cate.fit(X.primary = X[, 2, drop = FALSE],
                           X.nuis = X[, 1, drop = FALSE],
                           Y = Y, r = q, fa.method = "pc", adj.method = "nc",
                           nc = ctl, nc.var.correction = FALSE, calibrate = FALSE)

cout_post <- cate::cate.fit(X.primary = X[, 2, drop = FALSE],
                            X.nuis = X[, 1, drop = FALSE],
                            Y = Y, r = q, fa.method = "pc", adj.method = "nc",
                            nc = ctl, nc.var.correction = FALSE, calibrate = TRUE)

beta_t_med <- median(cout_pre$beta.t)
adjusted_t <- (cout_pre$beta.t - beta_t_med) / stats::mad(cout_pre$beta.t)

max(cout_post$beta.t - adjusted_t)
```

Note that to just inflate variance by MAD rather than by median, just do MAD of betahat^2 / sigmahat^2. So I am just going to say that we use

```{r}
cout_sigma2 <- cout_pre$beta.cov.row * diag(cout_pre$beta.cov.col) / n
max(cout_pre$beta.t - cout_pre$beta / sqrt(cout_sigma2))

## MAD inflation
stats::mad(cout_pre$beta^2 / cout_sigma2, center = 0)

## Control Gene Inflation
mean(cout_pre$beta[ctl]^2 / cout_sigma2[ctl])
```


Does `cout_pre` contain the unadjusted standard errors from `vruv4`

It does, though cate uses a different multiplicative factor. Basically CATE does not account for the degrees of freedom.

```{r}
vout_gls <- vicar::vruv4(Y = Y, X = X, ctl = ctl, k = q, cov_of_interest = 2,
                         likelihood = "normal", limmashrink = FALSE, include_intercept = FALSE)
vout_ols <- vicar::vruv4(Y = Y, X = X, ctl = ctl, k = q, cov_of_interest = 2,
                         likelihood = "normal", limmashrink = FALSE, gls = FALSE,
                         include_intercept = FALSE)
```

First, `ruv::RUV4` returns the same result as `vicar::vruv4` with `gls` set to `FALSE`.

```{r}
plot(c(vout_ols$sebetahat_ols) ^ 2, ruv4out$sigma2 * ruv4out$multiplier, 
     xlab = "vicar::vruv4 Var", ylab = "ruv::RUV4 VAR")
abline(0, 1)
```

Now, the variances of `cate::cate` and `vicar::vruv4` are off by a multiplicative constant.

```{r}
plot(c(vout_gls$sebetahat_ols ^ 2), cout_sigma2)
abline(0, 1)
```

`cate`'s `beta.cov.row` is the same as `vicar`'s `sigma2` --- but scaled differently. Specifically, `vicar` divideds by `n - k - q`, while `cate` divides just by `n - k`.
```{r}
vout_gls$sigma2 / cout_pre$beta.cov.row * (n - k - q) / (n - k)
```

`Gamma` in `cate` is the same as `alphahat` in `vicar`
```{r}
max(cout_pre$Gamma - vout_gls$alphahat)
```

`Z2` in `vicar` and `alpha` in `cate` differ only by a constant factor
```{r}
vout_gls$Z2 / cout_pre$alpha
O <- t(qr.Q(qr(X), complete = TRUE))
U <- O[1:2, ] %*% X
U11 <- U[2:2, 2:2, drop = FALSE]
U11 ## Just the last lower-triangular element in the R of the QR decomposition of X
```

Here is how `cate` calculates `beta.cov.col`
```{r}
cout_pre$beta.cov.col
solve(crossprod(X) / n)[2, 2] + crossprod(vout_gls$Z2) / U11 ^ 2 ## U11 is lower term in lower triangular matrix of QR decomposition of X.
```

We can get equivalence between `cate` and `vicar` by changing `n` above to `n-k`.

```{r}
XZ <- cbind(X, vout_gls$Zhat)
mult_val <- solve(crossprod(XZ))[2,2]
mult_val
(solve(crossprod(X) / (n - 2))[2, 2] + crossprod(vout_gls$Z2) / U11 ^ 2 ) / (n - 2) ## Divide by n-k instead of n to get equivalence.
```

Asymptotically, these variance estimates are the same, but can differ bigly in when the sample size is small.

```{r}
sessionInfo()
```
